Raising the Role of Vocabulary Hubs for Semantic Data Interoperability in Dataspaces

### 2024-10-01 EBDVF-2024/EBDVF-2024-presentation/presentation.qmd

---
title: "Raising the Role of Vocabulary Hubs for Semantic Data Interoperability in Dataspaces"
author: "Robert David, Petar Ivanov, Vladimir Alexiev"
date: 01-Oct-2024
subtitle: "3rd Workshop on Semantic Interoperability in Data Spaces"
format: revealjs
metadata-files:
  - ../_metadata.yml
---
# Acknowledgements
<div class="centered-text">
<img src="img/swc-logo-web.svg" width="45%">

<img src="img/Ontotext-Logo-vector.svg" width="45%">

<img src="img/underpin_logo_horizontal_icon_title_tagline_color_light_bg.svg" width="45%">

@@

- the european data economy depends on the availability of data 
- technological foundation + services and tools to make use of it
- different industries and communities
- machine-learning systems need high-quality training data
    - predictive maintenance

@@

- interoperability is key for making use of shared data
- syntactic data exchange already solved, I.e. defined data formats and exchange standards
- semantic interoperability still an open challenge
- Starting point: IDS RAM and IM defines semantic metadata
- we need to extend it to cover **semantic interoperability for data**

# Approach
- Semantic Interoperability via a **Semantic Layer** for Dataspaces
    - based on IDS Vocabulary Hub
    - based on knowledge graph (KG) technologies
    - implemented by **GraphDB + PoolParty** products
- services + data for semantic interoperability
- data discovery & harmonisation
- semantic metadata + data for discovery and integration

# Use Cases
2 running projects:
DataBri-X
- energy community simulation and prediction
- legal knowledge graph + document analysis

UNDERPIN
- refinery predictive maintenance
- wind farms predictive maintenance

@@

Semantic Layer supports structured and unstructured data interlinked in the KG.
Structured
- Ontology-Based Data Access (OBDA)
- RDB to RDF Mapping Language (R2RML)
- CSV on the Web (CSVW)

Unstructured
- semantic document annotations via KG entities

@@

- reasoning based on W3C recommendations OWL & SHACL
- inference tagging expands semantic annotations in documents
- vocabulary crosswalks connect different similar vocabularies
- metadata inference to determine specific vocabularies for domains

@@

- harmonized and semantically integrated data from disparate providers
- richer metadata descriptions
- incoming data bound dynamically to semantic descriptions
- richer discoverability / easier data querying
- vocabulary crosswalks: expand queries and annotations by traversing the interlinked graph

@@

***Future work:***

>>>>

---
title: "Raising the Role of Vocabulary Hubs for Semantic Data Interoperability in Dataspaces"
author: "Robert David, Petar Ivanov, Vladimir Alexiev"
date: 01-Oct-2024
subtitle: "3rd Workshop on Semantic Interoperability in Data Spaces"
format: revealjs
metadata-files:
  - _metadata.yml
---

@@

- The European data economy depends on the availability of data 
- Technological foundation + services and tools to make use of it
- Different industries and communities
- Machine-learning systems need high-quality training data
    - predictive maintenance

@@

- Interoperability is key for making use of shared data
- Syntactic data exchange already solved, I.e. defined data formats and exchange standards
- Semantic interoperability still an open challenge
- Starting point: IDS RAM and IM defines semantic metadata
- we need to extend it to cover **semantic interoperability for data**

# Approach
- Semantic Interoperability via a **Semantic Layer** for Dataspaces
    - based on IDS Vocabulary Hub
    - based on knowledge graph (KG) technologies
    - implemented by **GraphDB + PoolParty** products
- Services + data for semantic interoperability
- data discovery & harmonisation
- Semantic metadata + data for discovery and integration

# Use Cases
Two running projects:

**DataBri-X**

- energy community simulation and prediction
- legal knowledge graph + document analysis

**UNDERPIN**

- refinery predictive maintenance
- wind farms predictive maintenance

@@

Semantic Layer supports structured and unstructured data interlinked in the KG.

**Structured**

::::{.columns}
::: {.column width="40%"}
- Ontology-Based Data Access
- RDB to RDF Mapping Language
- CSV on the Web (CSVW)
:::
::: {.column width="60%"}
![](./img/skos.png){width=15%}
![](./img/r2rml.png){width=15%}
![](./img/csv.png){width=15%}
:::
::::

**Unstructured**

- Semantic document annotations via KG entities

@@

- Reasoning based on W3C recommendations OWL & SHACL
- Inference tagging expands semantic annotations in documents
- Vocabulary crosswalks connect different similar vocabularies
- Metadata inference to determine specific vocabularies for domains

@@

- Harmonized and semantically integrated data from disparate providers
- Richer metadata descriptions
- Incoming data bound dynamically to semantic descriptions
- Richer discoverability / easier data querying
- Vocabulary crosswalks: expand queries and annotations by traversing the interlinked graph

@@

**Future work:**
